\textls[-0]{\thetitle}

### 2023-03-24 metatheory2.tex
Phrasing.

Unlike in \cref{sec:metatheory2}, which uses named binders, in this appendix we
use \citeauthor{debruijn1964} indices and \emph{explicit
substitutions}~\cite{abadi1989subst,substcalculus}:
substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}, which
applies~$\gm$ to the type~$T$) are part of the syntax rather than
meta-operations. We specifically use based path induction
as eliminator for identity type~\cite{program2013homotopy}.
We work in an intrinsically typed setting: terms are well typed by construction, 
%
and consequently some premises are omitted without ambiguity and can be inferred from the context. \EDJ{I added this sentence back here. Why removed this sentence earlier?}\YZreply{I don't find a strong causal connection between the two parts of the sentence}
% 
This style of syntax formulation follows a recent
trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}
known as the ``algebraic presentation'' of MLTT.
%We introduce judgmental equality ($\equiv$)
%for derivations (of all kinds of judgements). In other words, we will have judgemental equality between terms, 
%contexts, explicits substitutions, and etc. 
%This style of syntax formulation has been carried out by the recent trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019} dubbed as ``algebraic presentation'' of MLTT. 
Moreover, universe levels are explicit in this appendix.

Other aspects of the syntax remain the same as in \cref{sec:metatheory2}.
In particular,
%following \citet{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019},
the syntax is still considered as being quotiented by the judgmental equalities.
Quotienting facilitates coercion along equalities.
Furthermore, the semantic model we develop in this appendix respects these
judgmental equalities by construction.
%This means 
%(1) our model have to respect the judgemental equalities to be a legitimate model for the quotiented syntax, and (2) we have convertibility for free---when we have definitionally equal $A \equiv B$, thanks to the quotient syntax, the derivation of $t : A$ is the same as the derivation of $t : B$.

@@

Consequently, function application changes to a equivalent formulation, and becomes a ``direct inverse'' of typing rule for function abstraction. For example, the named notation $\napp{f}{t}$ can be equivalently represented by $\sub{\app{f}}{\SubstExt{\SubstId}{t}}$.

Finally, we have \ruleref{sub/dbj/shift} defined using \ruleref{sub/wk} and \ruleref{tm/var}. This rule applying substitution $\gm$ on the earlier part of the context. We usually omit $A$ in the $\gm^{\uparrow A}$ because it can be inferred from the context.

>>>>

Unlike in \cref{sec:metatheory2}, which uses named binders, in this appendix we
use \citeauthor{debruijn1964} indices and \emph{explicit
substitutions}~\cite{abadi1989subst,substcalculus}:
substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}, which
applies~$\gm$ to the type~$T$) are part of the syntax rather than
meta-operations.
We work in an intrinsically typed setting: terms are well typed by construction, 
%
and consequently some premises are omitted without ambiguity and can be inferred from the context. \EDJ{I added this sentence back here. Why removed this sentence earlier?}\YZreply{I don't find a strong causal connection between the two parts of the sentence}
% 
This style of syntax formulation follows a recent
trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}
known as the ``algebraic presentation'' of MLTT.
%We introduce judgmental equality ($\equiv$)
%for derivations (of all kinds of judgements). In other words, we will have judgemental equality between terms, 
%contexts, explicits substitutions, and etc. 
%This style of syntax formulation has been carried out by the recent trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019} dubbed as ``algebraic presentation'' of MLTT. 
Moreover, universe levels are explicit in this appendix.

Other aspects of the syntax remain the same as in \cref{sec:metatheory2}.
In particular,
%following \citet{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019},
the syntax is still considered as being quotiented by the judgmental equalities.
Quotienting facilitates coercion along equalities.
Furthermore, the semantic model we develop in this appendix respects these
judgmental equalities by construction.
%This means 
%(1) our model have to respect the judgemental equalities to be a legitimate model for the quotiented syntax, and (2) we have convertibility for free---when we have definitionally equal $A \equiv B$, thanks to the quotient syntax, the derivation of $t : A$ is the same as the derivation of $t : B$.
We use based path induction
as the elimination principle for identity types~\cite{program2013homotopy}.

@@

Consequently, function application changes to an equivalent formulation, and becomes a ``direct inverse'' of typing rule for function abstraction. For example, the named notation $\napp{f}{t}$ can be equivalently represented by $\sub{\app{f}}{\SubstExt{\SubstId}{t}}$.

Finally, we have \ruleref{sub/dbj/shift} defined using \ruleref{sub/wk} and \ruleref{tm/var}. This rule applying substitution $\gm$ to the earlier portion of the context. We usually omit $A$ in the $\gm^{\uparrow A}$ because it can be inferred from the context.
